Project 1 Checkpoint

### 2015-02-27 project1/checkpoint/main.tex
Adding average car length
Adding a summary section a few citations

We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.

@@

The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
CoffeeScript removes some of the boilerplate associated with setting up prototype chains and other object-oriented
concepts in JavaScript.
Backbone \cite{backbone} along with Underscore \cite{underscore} and jQuery \cite{jquery} were used to provide some
application structure, like basic eventing, and the sorted
collections that are used to form the basis for the event queue and \texttt{IntersectionQueue}. Lastly, Browserify
\cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.

\section{Current Build and Running Instructions}
The simulation is currently deployed at \cite{aprilandchip} and currently runs a fixed number of cars through stepping
at 1 second real time being equal to 100ms. Eventually, this will be adjustable in real-time, and better visualized.
Cars are generated regularly at 10 (+ an uniformly random number between 0 and 1) second intervals. Light timings are
currently hard-coded as well. Currently, we're only modeling a single direction, and ignoring turns, and variations
between cars.

To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
simulation.

>>>>

We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
have the same length. These values can be calculated from the NGSIM data set \cite{ngsim} (which yields an average car length of 16.407).

@@

The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
CoffeeScript removes some of the boilerplate associated with setting up prototype chains and other object-oriented
concepts in JavaScript.
Backbone \cite{backbone} along with Underscore \cite{underscore} and jQuery \cite{jquery} were used to provide some
application structure, like basic eventing, and the sorted
collections that are used to form the basis for the event queue and \texttt{IntersectionQueue}. Lastly, Browserify
\cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.
Graphing is performed via inline SVG \cite{svg}.

\section{Current Build and Running Instructions}
The simulation is currently deployed at \cite{aprilandchip} and provides a UI that allows the light change intervals to
be updated in real time. Moreover, the simulation speed can be adjusted in real-time as well. Worth noting in the UI is
a log of all events that are passing through the system. This is live filterable. Also, a visualization of the current
signal state, along with the current wait times for the first 5 cars in line is provided. Statistics are provided in
real-time, providing both the observed input distribution of car inter-arrival times, and the output distribution of
car wait times.

To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
simulation. To rerun the simulation, simply refresh the browser

\section{Results}
Since light timing is adjustable in real-time, the actual average wait times can vary (even disregarding randomness).
However, a few general trends tend to appear. First off, the distribution of wait times is bimodal. This can be
somewhat observed in the NGSIM data set. This appears to be a result of the dichotomy of cars arriving when the light
is green, versus when it is red. Cars that arrive during a green light in general do not have to wait long at all
before they can cross the intersection. However, cars that arrive during a red light will be forced to wait, often
behind other cars. This causes their wait time to be much higher since they actually had to queue up before they could
exit the system. Unsurprisingly, most cars that catch the red light end up waiting around the duration of red time.

A second observation is the system saturation. There is a point where the light timing is so terrible that the system
cannot push cars through as fast as they come. In queueing theory terms, the system is referred to as being
``oversaturated``. After tweaking some values, this can be triggered by setting the red duration to 60 seconds and
slowing adjusting the yellow and green down (my testing resulted in oversaturation around yellow and green with 1 and
25 second durations respectively).

\section{Conclusion}
